---
title: "Week 2 Workbook Solutions"
description: |
author:
  - name: Joseph Bulbulia
    url: https://josephbulbulia.netlify.app
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0002-5861-2056
output:
  distill::distill_article:
    self_contained: false
    toc: true
    highlight: haddock
    highlight_downlit: false
---


### Create Repos (if you have not already)

  1. Create a repository for 447-journals (if you don't have one already)
  2. Create a repository for your 447-workbooks. 
  3. You should initialise a Readme statement and a .gitignore document (make sure to click the R .gitignore) 


### Solution

Students should confirm that they've done this


### Create R projects

  1. Create R-projects in each of these repositories.
  2. In your workbooks folder, using the `+ New Folder` command, create a folder called `data` (or similar) This is where you'll store your data.
  4. Create a file called `figs` (where you'll stor your figures)
  5 Create a folder called `workbooks`, or similar this is where you'll store your weekly workbooks. 

### Solution

If students did not create a data folder, then the `here` package task won't work.
  
### Workbook 2

 
### Q1 Open an R markdown document, and filling in your name

### Q2. The first code chunk is for your preferences. 


```{r setup, include=TRUE, eval = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Q3. Find the keyboard shortcuts menu. 

### Q4. Memorise the keyboard shortcut for insert code chunk which is: ⌥⌘I

### Q5. Use the mac to insert a code chunk. 

### Q6. Memorise the shortcut for evaluate which is ⌘-return^  use for all tasks in this workbook. 

### Q7. Use R to do 10 mathematical calculations 

### Solution 

Students should confirm that they've done this, e.g.:
  
```{r add1 }
4 + 5
```

```{r add2 }
3 ^ 3
```

Show your work 


### Q8. Using R, find the square root of 324

### Solution 

Students should show their works*

*JB: Students should confirm that they've done this*

  9. In the `iris` dataset, find the flower with the longest Sepal.Length. (hint, use the `max` function.)  Show the column details for this flower only. 
  
  
```{r maxiris}
# answer:
max(iris$Sepal.Length)
```

  
  10. In the `iris` dataset, find the flower with the shortest Sepal.Length. (hint, use the `min` function.)
  

```{r miniris}
# answer:
min(iris$Sepal.Length)
```


  11. Calculate the difference bewteen the flower with the longest and shortest Sepal.Length
  
```{r diffirise}
# answers
max(iris$Sepal.Length) - min(iris$Sepal.Length)

# or
a <- max(iris$Sepal.Length)
b <- min(iris$Sepal.Length)
```

  12. Make a 7 column  x  100 row dataframe using `c`, `:`, `dataframe`

```{r sevendf}
# some quick fixes
df <- data.frame(
  col1 = sample(1000, 100, replace = TRUE),
  col2 = rnorm(n = 100, mean = 0, sd = 1),
  col3 = 1:100,
  col4 = rbinom(n = 100, size = 1, prob = .5),
  col5 = rpois(n = 100, lambda = 10),
  col6 = sample(c("a", "b", "c", "d"), 100, replace = TRUE),
  col7 =  rep(1:10, 10)
)
```



  13. Rename the columns of the dataframe using the names of snow white dwarfs, or any names.

```{r renameseven}
# using base R
names(df)[] <-
  c(
    "sleepy",
    "grumpy",
    "happy",
    "smelly",
    "lovely",
    "daggy",
    "crazy"
    )
```

  14. Using the dataset `woman`, write a linear model for `height ~ weight`
  
```{r lmwomen}
# basic linear model

m1 <-  lm ( height ~ weight, data = women)
```

  15. Create a table and coefficient plot for this model
  
```{r plotwomen}
# coefficient plot
sjPlot::plot_model(m1)

# table
sjPlot::tab_model(m1)
```

  16. Using `ggeffects`, create a prediction plot for this model.

```{r predictwomen}
plot(
  ggeffects::ggpredict( m1 , effects = "height"),
  add.data = TRUE
)
```


<aside>
Next week we'll be creating plots like this
```{r smoothgraph}
library(ggplot2)
ggplot2::ggplot(data = women) +
  geom_point(mapping = aes(x = height, y = weight)) +
  geom_smooth(method = loess, aes(x = height, y = weight)) +
  theme_classic() 
```
</aside>


  17.  Explain what is being calculated here

```{r proportionwomen}
# this is a proportion (i.e. the number of instances / the total number of cases)
sum(women$weight > 140) / length(women$weight)
```
   
  18. Calculate the proportion of women who are over the mean weight of women

```{r calproportionwomen}
# This is onle approach
# Find the mean woman weight
mw <- mean( women$weight )

# calculae the proportion of women who are over that weight 
sum(women$weight > mw) / length(women$weight)
```


  19. What are the advantages and disadvantages of creating more breaks in the Petal.Length indicator?  Clarify your answer in a paragraph. 

```{r chunkyhist}
# advantates: major pattern of data evident; clearly two clusters (at least); no tempatation to over interpret patters in the graph

# disadvantages, looks as if there is a cluster at zero; only two modes in the dataset are discernable
hist( iris$Petal.Length )
```

```{r finhist}
# advantages: we can find more than one mode;  distribution not clustered at zero
# disadvantages: with such a small sample, it is tempting to read too much into all the modes. 

hist( iris$Petal.Length, 
      breaks = 100 )
```

  20. Spot the error in this code
  
```{r errorincode, eval = FALSE, echo = TRUE}
# here is one method. 
mh <- mean (women$height)
sum(women$weight > mh) / length(women$height)


# should be `sum(women$height > mh)
```

  21. Reorder the columns of the `woman` dataset so that weight comes before height. Then rename the columns "w" and "h". 

```{r reordercols, eval = TRUE, echo = TRUE}
# here is one method
# Bind columns as data frame
dfa <- cbind.data.frame(women$weight, women$height)
# change names
names(dfa)[] <- c("w", "h")

```

  22.  Read data into R using the following method:
  

```{r newdata, echo = FALSE, eval = FALSE}
# g-bayes making data (internal)

# Note that this won't work without a `data` folder
library(tidyverse)
n <- 100
yourtibble <- tibble(
  id = 1:n,
  weight = rnorm(n, 80, 20),
  height = rnorm(n, (weight + 10) * 1.8, 10)
)
plot(height ~ weight, yourtibble)
testdata1 <- data.frame(yourtibble)
write_csv(testdata1, here::here("data", "testdata1.csv"))
```


```{r readata, eval = TRUE, echo = TRUE}
# read data from file
library("readr")
testdata<- readr::read_csv(url("https://raw.githubusercontent.com/go-bayes/psych-447/main/data/testdata1.csv"))
str(testdata)

```

   Save data into your data folder using the following method
   
```{r herepackage, eval=FALSE, echo = TRUE}
library("here")
# save data using here
saveRDS(testdata, here::here("data", "td.RDS"))
```
   
   Read data back into R
   
```{r readdata, eval = FALSE}
td <- readRDS(here::here("data", "td.RDS"))
```

  23. Using the `td` dataset, write a linear model for `height ~ weight` as above

```{r secondlm, echo = FALSE, eval = FALSE}
m1<-lm(height ~ weight, data = td)
summary(m1)
plot(height ~ weight, data = td)
```

```{r secondreaddata}
td <- readRDS(here::here("data", "td.RDS"))
# note a clever student will change the object to something other than `m1` which we used above
m2<-lm(height ~ weight, data = td)

sjPlot::tab_model(m2)
```

  24. Create a coefficient plot
  
```{r plotsecond}
sjPlot::plot_model(m2)
```



  25. Create a prediction plot
  

```{r predictsecond}
plot(
  ggeffects::ggpredict( m2 , effects = "height" ),
  add.data = TRUE
)
```
  
Extra credit, **how would you interpret the intercept in this model?***


```{r interceptinterpretation}
# a zero intercept makes no sense because biologically speaking weight can't be zero
library( ggplot2 )
summary( m2 )
```

```



  
  




