---
title: "Workbook 6"
output: 
   learnr::tutorial:
       progressive: TRUE
       allow_skip: TRUE
       ace_theme: "iplastic"
       df_print: "kable"
       highlight: "kate"
runtime: shiny_prerendered
description: "Workbook 6."
---

```{r setup, include = FALSE}
library('learnr')
gradethis::gradethis_setup()
knitr::opts_chunk$set(error = TRUE)
learnr:::ACE_THEMES #for theme
```

## Warm Up


### W1.

Simulate 25 draws from a normal distribution with a mean of 10 and and a standard deviation of 2.

```{r rnorm1, exercise = TRUE}
set.seed(123)
rnorm(n = , mean = , sd = )
```

```{r rnorm1-solution}
set.seed(123)
rnorm(n = 25, mean = 10, sd = 2)
```

```{r rnorm1-code-check}
grade_code()
```



What does `f(n)` evaluate to, for n = 10? (See if you can work it out in your head)



### W2

Below is code for creating a sample of 100 with means of 0, 10, and 50.  Assign standard deviations of 1, 2, and 5 to each the the three means

```{r rnorm2, exercise = TRUE}
set.seed(123)
rnorm(n = 12, mean = c(0, 10, 50)  )
```


```{r rnorm2-solution}
set.seed(123)
rnorm(n = 12, mean = c(0, 10, 50) , sd = c(1,2,5) )
```

```{r rnorm2-code-check}
grade_code()
```

### W3


Create two samples of 100, x and y, with N = 100 and with a mean of 0 and an sd of 1.  Use `plot()` to quickly graph the relationship between this two distributions 

```{r plotrnorm, exercise = TRUE}
x = rnorm(n = , mean = , sd = )
y = rnorm(n = , mean = , sd = )
plot(y ~ x)
```

```{r plotrnorm-solution}
x = rnorm(n = 100, mean = 0, sd = 1)
y = rnorm(n = 100, mean = 0, sd = 1)
plot(y ~ x)
```

```{r plotrnorm-code-check}
grade_code()
```


Repeat this code and inspect the output. How is the graph changing? 


### W4

Create a uniform distribution by producing 10 numbers between between 0 and 1 using the `runif()` command.  

```{r runif1, exercise = TRUE}

```

<div id="filter-hint">
**Hint:** You may want to use `min = ` and `max =` 
</div>


```{r runif1-solution}
runif(n = 5, min = 0, max = 1)
```

```{r runif1-code-check}
grade_code()
```

### W5

Sample 100 draws from a normal distribution with mean 5 and sd = 1 and call the vector 'y',  Sample 100 draws from a uniform distribution from 1 to 50 and call the vector 'x'.  Make a dataframe called `df` and regress y on x. 

```{r dfrg, exercise = TRUE}

```

<div id="filter-hint">
**Hint:** `lm( this ~ that)`
</div>


```{r dfrg-solution}
set.seed(16)
y = rnorm(n = 100, mean = 5, sd = 1)
x = runif(n = 100, min = 1, max = 50)
df = data.frame(y,x)
lm(y ~ x, data = df)
```

```{r dfrg-code-check}
grade_code()
```


### W6 

Letters

We can create letters like this

Here are the first two `letters` of the English alphabet.

```{r letters, exercise = TRUE}
letters[1:2]
```

And the final 15 `LETTERS` of the English alphabet.

```{r LETTERS, exercise = TRUE}
LETTERS[12:26]
```

Here is how we use each to create a vector of letters


```{r each, exercise = TRUE}
rep(x = letters[1:2], each = 3)
```


We can create a two-group vector again, how can we get the sequence to alternate three times, instead of repeat three times? 

```{r times, exercise = TRUE}
rep(letters[1:2], )
```

<div id="filter-hint">
**Hint:** `times` they are a chang'n
</div>


```{r times-solution}
rep(letters[1:2], times = 3)
```


```{r times-code-check}
grade_code()
```

### W7 

Make a two-group variable of length 5.  

```{r lengthout, exercise = TRUE}
rep( , length.out = )
```

```{r lengthout-solution}
rep(letters[1:2], length.out = 9)
```

Describe how the groups are unbalanced.  

W10

Repeat each of the two letters twice, 3 times. 

```{r times2, exercise = TRUE}
rep(letters[1:2],  )
```

```{r times2-solution}
rep(letters[1:2], times = c(2, 4) )
```

```{r times2-code-check}
grade_code()
```

### W8 

- Make a variable named `group` that repeats the first two `letters` 3 times each (element-wise).  
- Make a variable named `response` that samples from a normal distribution with mean zero and a standard deviation of 1. create a data frame

```{r dat1, exercise = TRUE}
group = 
response = 
data.frame(group,
           response)
```

```{r dat1-solution}
group = rep(letters[1:2], each = 3)
response = rnorm(n = 6, mean = 0, sd = 1)
data.frame(group,
           response)
```

```{r dat1-code-check}
grade_code()
```


Note you can create the data frame from the previous exercise at one go:

```{r dat2, exercise = TRUE}
data.frame(group = rep(letters[1:2], each = 3),
           response = rnorm(n = 6, mean = 0, sd = 1))
```


### W9

Recall that we can create big letters this way:


```{r dateach-solution}
rep(LETTERS[3:5], each = 2)
```


Imagine that the you have two factors `a` and `b` that are crossed in three groups say `C`,`D`,`E`. (e.g. male and not_male in manipulation_1, manipulation_2, and a control condition)

using the `times` function, create a dataframe in which the factors are crossed by the three conditions exactly once


```{r dat3, exercise = TRUE}
data.frame(group = rep(letters[1:2], each = 3),
           factor = ,
           response =  rnorm(n = 6, mean = 0, sd = 1) )
```


```{r dat3-solution}
data.frame(group = rep(letters[1:2], each = 3),
           factor = rep(LETTERS[3:5], times = 2),
           response = rnorm(n = 6, mean = 0, sd = 1) )
```

```{r dat3-code-check}
grade_code()
```


Next create a dataframe with the same properties using the `length.out` function

```{r dat4, exercise = TRUE}
data.frame(group = rep(letters[1:2], each = 3),
           factor = ,
           response =  rnorm(n = 6, mean = 0, sd = 1) )
```


```{r dat4-solution}
data.frame(group = rep(letters[1:2], each = 3),
           factor = rep(LETTERS[3:5], length.out = 2),
           response = rnorm(n = 6, mean = 0, sd = 1) )
```

```{r dat4-code-check}
grade_code()
```

### W10

Recall how rnorm works:

Create n = 6 responses divided into two groups, one with a mean of 3, and one with a mean of 2, each has a standard deviation of 1 and both sample from a normal distribution

```{r twomeans, exercise = TRUE}
rnorm(n = 6, mean = , sd = )
```

```{r twomeans-solution}
rnorm(n = 6, mean = c(5, 10), sd = 1)
```

```{r twomeans-code-check}
grade_code()
```

Using the code above, create a data frame in which there are two factors crossed by three groups and the means differ between the groups


```{r dat5, exercise = TRUE}
data.frame(group = rep(),
            response = rnorm() )
```


<div id="dat5-hint">
**Hint:** You need to repeat the whole vector three times instead of repeating the values elementwise.
</div>



```{r dat5-solution}
data.frame(group = rep(letters[1:2], times = 3),
            response = rnorm(n = 6, mean = c(5, 10), sd = 1) )
```


### FUN FUN FUN

The following was on twitter yesterday.  Can you imagine the solution in your head? Try

```{r function1, exercise = TRUE}
f <- function(x) {
  f <- function(x) {
    f <- function() {
      x ^ 2
    }
    f() + 1
  }
  f(x) * 2
}
f(n)
```


```{r function1-solution}
f <- function(x) {
  f <- function(x) {
    f <- function() {
      x ^ 2
    }
    f() + 1
  }
  (y) * 2
}
f(10)
```

## Work book

Your final exercises is a 2500 report for a substantive research question in psychological science. 

This week, write a two-three paragraph introduction to your report and summarise the sample information using graphical and tabular (and any other methods) that you have learned in this class. 

If you have not collected data, you may either see us for a dataset that might be suitable or alternatively you may simulate data. 

To keep the report streamlined, you may include an appendix. 

Reports should contain a bibliography 





### Acknowledgments

The warmup exercises are from Ariel Muldoon's tutorial on simulation, with only minor modifications: https://aosmith.rbind.io





