---
title: "Workbook 3"
description: |
author:
  - name: Joseph Bulbulia
    url: https://josephbulbulia.netlify.app
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0002-5861-2056
  - name: Johannes Karl 
    url: https://johannes-karl.com
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0001-5166-0728
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

### Libraries we will need
```{r}
library("tidyverse")
library("patchwork")
library("readr")
library("sjPlot")
```

```{r}
# set theme
theme_set(theme_classic())
```


### Question 1: Why is this graph not printing any output? 

```{r}
library("tidyverse")
ggplot(data = mtcars) + 
  aes(mpg, wt, colour=factor(cyl))
```



### Question 2. Using the mpg dataset, graph the relationship between city milage and highway mileage by year manufacture



### Question 3. Edit this graph so that the x axis and the y axis both start at 0


```{r}
# Create graph and add title
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  labs(title = "Relationship bewtween engine displacement and fuel efficiency in the mpg automobile dataset") + 
  xlab("Engine displacement in (units)") + 
  ylab("Highway miles per liter")
```




### Question 4: what is one benefit and one limitation for this graph above (in which the x and y values start at 0?)




### Quetion 5.  Which of these two graphs do you prefer and why? 

```{r}
g1 <-ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour =  class )) + 
  labs(title = "Relationship bewtween engine displacement and fuel efficiency in the mpg automobile dataset") + 
  xlab("Engine displacement in (units)") + 
  ylab("Highway miles per liter") 

```



```{r}
g2 <-ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape =  class )) + 
  labs(title = "Relationship bewtween engine displacement and fuel efficiency in the mpg automobile dataset") + 
  xlab("Engine displacement in (units)") + 
  ylab("Highway miles per liter") 
```
  
  
  
```{r layout="l-body-outset", fig.width=12, fig.height=12}
library("patchwork")

g1 / g2 + plot_annotation(title = "Which plot do you prefer and why?", tag_levels = 'a')
```



### Quetion 6. add a facet to this graph for the "class" variable

```{r}
g2 <-ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape =  class )) + 
  labs(title = "Relationship bewtween engine displacement and fuel efficiency in the mpg automobile dataset") + 
  xlab("Engine displacement in (units)") + 
  ylab("Highway miles per liter") 
```


### Question 7.  which graph is more informative and why?


```{r  dualfig, layout="l-body-outset", fig.width=12, fig.height=12, echo = FALSE}
g3 <- ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

g4 <- ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  facet_wrap( ~ class, nrow = 2)

g3 / g4 + plot_annotation("Which graph is more informative and why?", tag_levels = 'a')
```





### Question 8.  remove the legend from the facet graph above (g4)


```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  facet_wrap( ~ class, nrow = 2)
```



### Question  9  Convert the y variable to "numeric" and graph the relationship betweeen `religiousity` (x-axis) and `thr_mus`` (y-axis) in the ISSP dataset. Create new axis labels

Download the ISSP questionaire used in this study [here]: 
(https://github.com/go-bayes/psych-447/blob/main/data_raw/ISSP/ISSP_2018_Religion_Questionnaire_final_version1-2.pdf)

Note^[Data were collected with Barry Milne and Martin Van dataset; the data are only authorised for the purposes of teaching.]


```{r}
# read the issp dataset for questionaire see: ISSP_2018_Religion_Questionnaire_final_version1-2.pdf
# 
# subset of data from the issp dataset
issp <- readr::read_csv2(url("https://raw.githubusercontent.com/go-bayes/psych-447/main/data/issp.csv"))
```


```{r eval=FALSE}
# note that we need to check our data
head(issp)

# when we do we find that classes of the variables need to be adjusted
str(issp)
```

Let's get the data into shape using dplyr. Run this code below

```{r}
ip <- issp %>%
  mutate(
    id = factor(id),
    thr_ath = as.factor(thr_ath),
    thr_bd = as.factor(thr_bd),
    thr_ch = as.factor(thr_ch),
    thr_hd = as.factor(thr_hd),
    thr_jw = as.factor(thr_jw),
    thr_ms = as.factor(thr_ms),
    neg_ath = as.factor(neg_ath),
    neg_bd = as.factor(neg_bd),
    neg_ch = as.factor(neg_ch),
    neg_hd  = as.factor(neg_hd),
    neg_jw = as.factor(neg_jw),
    neg_ms = as.factor(neg_ms),
    wave  = as.factor(wave),
    nzeuro = as.factor(nzeuro),
    eduyears = as.numeric(eduyears),
    male = as.factor(male),
    age = as.numeric(age),
    rightwing = as.numeric(rightwing),
    rural = as.factor(rural),
    religiosity = as.numeric(religiosity)
  )
```





### Question 10.  Note that I have  the following graph should start from 1 and run to 4 but currently runs from 0-4.  Fix the graph 

```{r}
library(ggplot2)
ggplot(data = ip, aes(y = as.numeric(thr_ms), x = religiosity, colour = wave))  +  geom_jitter(alpha = .1) + 
  geom_smooth(method = lm, fullrange = FALSE, alpha = 0.1) +
   scale_y_continuous(limits = c(0,4))
```





### Extra Question 11. Find one way of improving the the following code and explain your answer


```{r}
library(sjPlot)
plot_xtab(
    ip$thr_ms,
    ip$wave,
    show.total = F,
    show.n = F,
    geom.colors = c("lightgreen", "darkred")
  ) +
  xlab("Threatened by Muslims") +  ylab("Frequency") +
  #scale_y_continuous(limits=c(0,7)) + #theme(plot.title = element_text(size=09))
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

```






