---
title: "Week 9 workbook and solutions"
description: |
author:
  - name: Joseph Bulbulia
    url: https://josephbulbulia.netlify.app
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0002-5861-2056
output:
  distill::distill_article:
    self_contained: false
    toc: true
    highlight: kate
    code_folding: true
---

Import the jittered NZAVS dataset


```{r setup, include=FALSE}
# setup
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  layout = "l-body-outset",
  fig.width= 12,
  fig.height= 10,
  collapse =TRUE,
  R.options = list(width = 60)
)
```
```{r  libraries}
### Libraries
library("tidyverse")
library("patchwork")
library("lubridate")
library("kableExtra")
library("gtsummary")
library("lubridate")
library("equatiomatic")
library("ggdag")
library("brms")
library("rstan")
library("rstanarm")
library("bayesplot")
library("easystats")
library("kableExtra")
library("broom")
library("tidybayes")
library("bmlm")
# if (!require(tidyLPA)) {
#   install.packages("tidyLPA")
# }
# rstan options
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores ())
theme_set(theme_classic())

```{r  nzdata, cache=TRUE}
# read data
# read data


nz_0 <- as.data.frame(readr::read_csv2(
  url(
    "https://raw.githubusercontent.com/go-bayes/psych-447/main/data/nzj.csv"
  )
))

# to relevel kessler 6 variables
f <-
  c(
    "None Of The Time",
    "A Little Of The Time",
    "Some Of The Time",
    "Most Of The Time",
    "All Of The Time"
  )



# get data into shape
nz_cr <- nz_0 %>%
  dplyr::mutate_if(is.character, factor) %>%
  select(
    -c(
      SWB.Kessler01,
      SWB.Kessler02,
      SWB.Kessler03,
      SWB.Kessler04,
      SWB.Kessler05,
      SWB.Kessler06
    )
  ) %>%
  dplyr::mutate(Wave = as.factor(Wave)) %>%
  dplyr::mutate(FeelHopeless = forcats::fct_relevel(FeelHopeless, f)) %>%
  dplyr::mutate(FeelDepressed = forcats::fct_relevel(FeelDepressed, f)) %>%
  dplyr::mutate(FeelRestless = forcats::fct_relevel(FeelRestless, f)) %>%
  dplyr::mutate(EverythingIsEffort = forcats::fct_relevel(EverythingIsEffort, f)) %>%
  dplyr::mutate(FeelWorthless = forcats::fct_relevel(FeelWorthless, f)) %>%
  dplyr::mutate(FeelNervous = forcats::fct_relevel(FeelNervous, f)) %>%
  dplyr::mutate(Wave = as.factor(Wave)) %>%
  dplyr::mutate(male_id = as.factor(Male)) %>%
  dplyr::mutate(date = make_date(year = 2009, month = 6, day = 30) + TSCORE) %>%
  dplyr::mutate(
    FeelWorthless_int = as.integer(FeelWorthless),
    FeelNervous_int =  as.integer(FeelNervous),
    FeelHopeless_int =  as.integer(FeelHopeless),
    EverythingIsEffort_int =  as.integer(EverythingIsEffort),
    FeelRestless_int =  as.integer(FeelRestless),
    FeelDepressed_int =  as.integer(FeelDepressed),
    HLTH.Fatigue_int = as.integer(HLTH.Fatigue + 1)
  ) %>%
  dplyr::mutate(yearS = TSCORE - min(TSCORE, na.rm = TRUE)) %>%
  dplyr::mutate(KESSLER6sum = as.integer(KESSLER6sum))

## 2018 only data
nz2018 <- nz_cr %>%
  dplyr::filter(Wave == 2018)

## 2019 only timeline
# Relevel covid timeline 2019
ord_dates_class_2019_only <- c("PreCOVID",
                               "JanFeb",
                               "EarlyMarch",
                               "Lockdown",
                               "PostLockdown")

nz2019 <- nz_cr %>%
  dplyr::filter(YearMeasured == 1) %>%
  dplyr::filter(Wave == 2019) %>%
  dplyr::group_by(Id) %>%
  dplyr::ungroup(Id) %>%
  dplyr::mutate(Covid_Timeline_cr =
                  as.factor(ifelse(
                    TSCORE %in% 3896:3921,
                    # feb 29 - march 25th
                    "EarlyMarch",
                    ifelse(
                      TSCORE %in% 3922:3954,
                      "Lockdown",
                      #march 26- Mon 27 April 2020
                      ifelse(
                        TSCORE > 3954,
                        # after april 27th 20202
                        "PostLockdown",
                        ifelse(TSCORE %in% 3842:3895,
                               # jan 6 to feb 28
                               "JanFeb",
                               "PreCOVID")
                      )
                    )
                  ))) %>%
  dplyr::mutate(Covid_Timeline_cr = forcats::fct_relevel(Covid_Timeline_cr, ord_dates_class_2019_only))



## 2018/2019 data set  
ord_dates_class <- c("Baseline",
                     "PreCOVID",
                     "JanFeb",
                     "EarlyMarch",
                     "Lockdown",
                     "PostLockdown")

nzl <- nz_cr %>%
  dplyr::filter(YearMeasured == 1) %>%
  dplyr::filter(Wave == 2018 | Wave == 2019) %>%
  dplyr::group_by(Id) %>%
  dplyr::filter(n() > 1) %>%
  dplyr::filter(n() != 0) %>%
  dplyr::ungroup(Id) %>%
  dplyr::mutate(yearS = (TSCORE - min(TSCORE)/365)) %>%
  dplyr::mutate(WSCORE = as.factor(WSCORE)) %>%
  dplyr::mutate(Covid_Timeline =
                  as.factor(ifelse(
                    TSCORE %in% 3896:3921,
                    # feb 29 - march 25th
                    "EarlyMarch",
                    ifelse(
                      TSCORE %in% 3922:3954,
                      "Lockdown",
                      #march 26- Mon 27 April 2020
                      ifelse(
                        TSCORE > 3954,
                        # after april 27th 20202
                        "PostLockdown",
                        ifelse(
                          TSCORE %in% 3842:3895,
                          # jan 6 to feb 28
                          "JanFeb",
                          ifelse(TSCORE %in% 3665:3841 &
                                   Wave == 2019,
                                 "PreCOVID",
                                 "Baseline"  # 3672 TSCORE or  20 July 2019))))))))
                          )
                        )
                      )
                    ))))%>%
  dplyr::mutate(Covid_Timeline = forcats::fct_relevel(Covid_Timeline, ord_dates_class))
```

```{r}
dplyr::glimpse(nzl)
```

### Assessment

#### Information
Link to the NZAVS data dictionary is 
[here](https://dl.dropboxusercontent.com/s/but0pz9h4jdpy6f/NZAVS-Technical-Documents-e02-Data-Dictionary.xlsx?dl=0)
 
Link to questions only [here](https://dl.dropboxusercontent.com/s/cv256sbjal4ucvo/nzavs-data-dictionary.htm?dl=0)
 
#### Select any one of the following three tasks

1. Write brief report that in which you estimating demographic and ideological  predictors of attitudes to government surveillance `Issue.GovtSurveillance` and attitudes to `Issue.RegulateAI` in New Zealand in 2018


<!-- ```{r} -->
<!-- nz$Issue.RegulateAI -->
<!-- gv <- bf(Issue.GovtSurveillance ~ yearS +  (1|p|Id)) -->
<!-- ai <- bf(Issue.RegulateAI ~ yearS + (1|p|Id)) -->
<!-- fit2 <- brm(gv + ai, data = nz2018) -->
<!-- summary(fit2) -->
<!-- ``` -->


2. Write a brief report explaining the predictors of stability and change in employment security  (`Emp.JobSecure`) from 2018-2020 (Waves 10/11 of the the NZAVS Jitter dataset). 

<!-- ```{r} -->
<!-- library(splines) -->
<!-- ho<- lme4::lmer(Emp.JobSecure  ~ yearS + (1|Id), data = nzl) -->
<!-- plot(ggeffects::ggpredict(ho), add.data = TRUE) -->

<!-- ho2<- lme4::lmer(Emp.JobSecure  ~ yearS + Covid_Timeline + (1|Id), data = nzl) -->
<!-- plot(ggeffects::ggpredict(ho2), add.data = TRUE) -->
<!-- ``` -->


3. Did people gain weight during New Zealand's Covid Lockdown? 

#### Marking criteria

1. Clarity and organisation in the descriptive component of your work. 

2. Clarity and organisation in description of your regression model.

3. Accuracy and insight in the interpretation of your regression model.

