---
title: "Week 3 Lab"
description: |
author:
  - name: Joseph Bulbulia
    url: https://josephbulbulia.netlify.app
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0002-5861-2056
  - name: Johannes Karl 
    url: https://johannes-karl.com
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0001-5166-0728
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

### Libraries we will need
```{r}
library("tidyverse")
library("patchwork")
library("readr")
```



### Question Question: Why is this graph not printing any output? 

```{r}
library("tidyverse")
ggplot(data = mtcars) + 
  aes(mpg, wt, colour=factor(cyl))
```


### Answer, there are no layers.


### Question 2. Using the mpg dataset, graph the relationship between city milage and highway mileage by year manufacture


### Answer
```{r}
ggplot(mpg,aes(x = hwy, y = cty, color=year)) + geom_point()+geom_smooth(method = loess)
```


### Question 3. Edit this graph so that the x axis and the y axis both start at 0


```{r}
# Create graph and add title
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  labs(title = "Relationship bewtween engine displacement and fuel efficiency in the mpg automobile dataset") + 
  xlab("Engine displacement in (units)") + 
  ylab("Highway miles per liter")
```


###  Answer

```{r}
  ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  labs(title = "Relationship bewtween engine displacement and fuel efficiency in the mpg automobile dataset") + 
  xlab("Engine displacement in (units)") + 
  ylab("Highway miles per liter") + 
  expand_limits(x = 0, y = 0)
```


### Question 4: what is one benefits and one limitation of this graph (in which the x and y values start at 0?)


### Answer

Example benefits:

  1. We clearly see that engine displacement starts at just below
  2. We clearly understand that no highway mileage below 10 mpg doesn't exist.
  
Example weakness

  1. Even if the thresholds are not evident, we can already see the thresholds without starting at graphs at zero (i.e. there is no information)
  2. Zero engine displacement and zero highway miles per gallon are not physically meaningful concepts, so starting the graphs at these thresholds adds no new information. 
  

### Quetion 5.  Which of these two graphs do you prefer and why? 

```{r}
g1 <-ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour =  class )) + 
  labs(title = "Relationship bewtween engine displacement and fuel efficiency in the mpg automobile dataset") + 
  xlab("Engine displacement in (units)") + 
  ylab("Highway miles per liter") 

```



```{r}
g2 <-ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape =  class )) + 
  labs(title = "Relationship bewtween engine displacement and fuel efficiency in the mpg automobile dataset") + 
  xlab("Engine displacement in (units)") + 
  ylab("Highway miles per liter") 
```
  
```{r layout="l-body-outset", fig.width=12, fig.height=12}
library("patchwork")

g1 / g2 + plot_annotation(title = "Which plot do you prefer and why?", tag_levels = 'a')
```

### Answer

Arguably, patterns are easier to detect using the colour aesthetic but there are no hard and fast rules



### Quetion 6.  add a facets to this graph for the "class" variable

```{r}
g2 <-ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape =  class )) + 
  labs(title = "Relationship bewtween engine displacement and fuel efficiency in the mpg automobile dataset") + 
  xlab("Engine displacement in (units)") + 
  ylab("Highway miles per liter") 
```


### Answer
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class) + 
  labs(title = "Relationship bewtween engine displacement and fuel efficiency in the mpg automobile dataset") + 
  xlab("Engine displacement in (units)") + 
  ylab("Highway miles per liter") 
```


### Question 7.  which graph is more informative and why?


```{r  dualfig, layout="l-body-outset", fig.width=12, fig.height=12, echo = FALSE}
g3 <- ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

g4 <- ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  facet_wrap( ~ class, nrow = 2)

g3 / g4 + plot_annotation("Which graph is more informative and why?", tag_levels = 'a')
```


### Answer

Arguably the facets make the role of class more evident by making the class indicator more salient. Where possible,i t is a good idea to declutter your graph.   


#### Question 8.  remove the legend from the facet graph above (g4)


```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  facet_wrap( ~ class, nrow = 2)
```

### Answer

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  facet_wrap( ~ class, nrow = 2) +
  theme(legend.position = "none") 
```


### Question  9  Convert the y variable to "numeric" and graph the relationship betweeen `religiousity` (x-axis) and `thr_mus`` (y-axis) in the ISSP dataset. Create new axis labels

Download the ISSP questionaire used in this study here: 
https://github.com/go-bayes/psych-447/blob/main/data_raw/ISSP/ISSP_2018_Religion_Questionnaire_final_version1-2.pdf


```{r}
# read the issp dataset for questionaire see: ISSP_2018_Religion_Questionnaire_final_version1-2.pdf
# 
# subset of data from the issp dataset
issp <- readr::read_csv2(url("https://raw.githubusercontent.com/go-bayes/psych-447/main/data/issp.csv"))
```


```{r eval=FALSE}
# note that we need to check our data
head(issp)

# when we do we find that classes of the variables need to be adjusted
str(issp)
```

Let's get the data into shape using dplyr. Run this code below

```{r}
ip <- issp %>%
  mutate(
    id = factor(id),
    thr_ath = as.factor(thr_ath),
    thr_bd = as.factor(thr_bd),
    thr_ch = as.factor(thr_ch),
    thr_hd = as.factor(thr_hd),
    thr_jw = as.factor(thr_jw),
    thr_ms = as.factor(thr_ms),
    neg_ath = as.factor(neg_ath),
    neg_bd = as.factor(neg_bd),
    neg_ch = as.factor(neg_ch),
    neg_hd  = as.factor(neg_hd),
    neg_jw = as.factor(neg_jw),
    neg_ms = as.factor(neg_ms),
    wave  = as.factor(wave),
    nzeuro = as.factor(nzeuro),
    eduyears = as.numeric(eduyears),
    male = as.factor(male),
    age = as.numeric(age),
    rightwing = as.numeric(rightwing),
    rural = as.factor(rural),
    religiosity = as.numeric(religiosity)
  )
```


#### Solution 


```{r}
library(ggplot2)
ggplot(data = ip, aes(y = as.numeric(thr_ms), x = religiosity, colour = wave)) + 
  geom_smooth(method = lm, fullrange = FALSE, alpha = 0.1) + 
  labs(title = "Religiosity predict Muslim acceptance post-Christchurch shootings") + 
  xlab( x = "Level of Religiosity (scale 1-7) ") + ylab ("acceptance of Muslims (1-4)")
```


### 

```{r}
library(ggplot2)
ggplot(data = ip, aes(y = as.numeric(thr_ms), x = religiosity, colour = wave))  +  geom_jitter(alpha = .1) + 
  geom_smooth(method = lm, fullrange = FALSE, alpha = 0.1) +
   scale_y_continuous(limits = c(1,4))
```



```{r}
library(ggplot2)
ggplot(data = issp, aes(y = as.numeric(thr_ms), x = rightwing, colour = wave)) + 
  geom_smooth(method = lm, fullrange = FALSE, alpha = 0.1) +
  scale_y_continuous(limits = c(0,4))
```
