---
title: "Week 2 Workbook"
description: |
author:
  - name: Joseph Bulbulia
    url: https://josephbulbulia.netlify.app
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0002-5861-2056
output:
  distill::distill_article:
    self_contained: false
    toc: true
---


### Create Repos (if you have not already)

  1. Create a repository for 447-journals (if you don't have one already)
  2. Create a repository for your 447-workbooks. 
  3. You should initialise a Readme statement and a .gitignore document (make sure to click the R .gitignore) 


### Create R projects

  1. Create R-projects in each of these repositories.
  2. In your workbooks folder, using the `+ New Folder` command, create a folder called `data` (or similar) This is where you'll store your data.
  4. Create a file called `figs` (where you'll stor your figures)
  5 Create a folder called `workbooks`, or similar this is where you'll store your weekly workbooks. 
  
### Workbook 2

  1. Open an R markdown document, and filling in your name
  2. The first code chunk is for your preferences. 


```{r setup, include=TRUE, eval = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  3. Find the keyboard shortcuts menu. 
  4. Memorise the keyboard shortcut for insert code chunk which is: ⌥⌘I
  5. Use the mac to insert a code chunk. 
  6. Memorise the shortcut for evaluate which is ⌘-return^  use for all tasks in this workbook. 
  7. Use R to do 10 mathematical calculations 
  
```{r}
4+5
```

```{r}
3^3
```

Show your work 



  8. Using R, find the square root of 324
  9. In the `iris` dataset, find the flower with the longest Sepal.Length. (hint, use the `max` function.)  Show the column details for this flower only. 
  10. In the `iris` dataset, find the flower with the shortest Sepal.Length. (hint, use the `min` function.)
  11. Calculate the difference bewteen the flower with the longest and shortest Sepal.Length
  12. Make a 7 column  x  100 row dataframe using `c`, `:`, `dataframe`
  13. Rename the columns of the dataframe using the names of snow white dwarfs, or any names.
  14. Using the dataset `woman`, write a linear model for `height ~ weight`
  15. Create a table and coefficient plot for this model
  16. Using `ggeffects`, create a prediction plot for this model.

<aside>
Next week we'll be creating plots like this
```{r}
library(ggplot2)
ggplot2::ggplot(data = women) + 
  geom_point(mapping = aes(x = height, y = weight)) + geom_smooth(method = lm, aes(x = height, y = weight)) + theme_classic() 
```
</aside>


  17.  Explain what is being calculated here

```{r}
sum(women$weight > 140) / length(women$weight)
```
   
  18. Calculate the proportion of women who are over the mean weight of women

```{r eval = FALSE, echo = FALSE}
mw <- mean(women$weight)
sum(women$weight > mw) / length(women$weight)
```
  
```{r eval = FALSE, echo = FALSE}
mh <- mean(women$height)
sum(women$height > mh) / length(women$height)
```


  19. What are the advantages and disadvantages of creating more breaks in the Petal.Length indicator?  Clarify your answer in a paragraph. 

```{r}
hist(iris$Petal.Length)
```

```{r}
hist(iris$Petal.Length, breaks = 100)
```

  20. Spot the error in this code
  
```{r eval = FALSE, echo = TRUE}
# here is one method. 
mh <- mean(women$height)
sum(women$weight > mh) / length(women$height)
```

  21. Reorder the columns of the `woman` dataset so that weight comes before height. Then rename the columns "w" and "h". 

```{r, eval = FALSE, echo = FALSE}
# here is one method
# Bind columns as data frame
dfa <- cbind.data.frame(women$weight, women$height)
# change names
names(dfa)[] <- c("w", "h")

```

  22.  Read data into R using the following method:
  

```{r echo = FALSE, eval = FALSE}

# g-bayes making data (internal)
library(tidyverse)
n <- 100
yourtibble <-tibble(
  id = 1:n, 
  weight = rnorm(n, 80,20), 
  height = rnorm(n, (weight+10)*1.8, 10)
)
plot( height ~ weight, yourtibble)
testdata1<-data.frame(yourtibble)
write_csv(testdata1,here::here("data", "testdata1.csv"))
```


```{r eval = FALSE, echo = TRUE}
# read data from file
library("readr")
testdata<- readr::read_csv(url("https://raw.githubusercontent.com/go-bayes/psych-447/main/data/testdata1.csv"))
str(testdata)
```

   Save data into your data folder using the following method
   
```{r eval=FALSE, echo = TRUE}
library("here")
# save data using here
saveRDS(testdata, here::here("data", "td.RDS"))
```
   
   Read data back into R
   
```{r eval = FALSE}
td <- readRDS(here::here("data", "td.RDS"))
str(td)
```

  23. Using the `td` dataset, write a linear model for `height ~ weight` as above

```{r echo = FALSE, eval = FALSE}
m1<-lm(height ~ weight, data = td)
summary(m1)
plot(height ~ weight, data = td)
```

  24. Create a coefficient plot
  25. Create a prediction plot
  
Extra credit, **how would you interpret the intercept in this model?***



  
  




