---
title: "Week 3 Lab"
description: |
author:
  - name: Joseph Bulbulia
    url: https://josephbulbulia.netlify.app
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0002-5861-2056
  - name: Johannes Karl 
    url: https://johannes-karl.com
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0001-5166-0728
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

## Setup

### Libraries

```{r}
library(tidyverse)
library(lubridate)
library(pmdplyr)
library(correlation)
library(ggraph)

```

Run this code
```{r eval=FALSE}
#easystats::install_easystats_latest()
```


### Import  data
```{r}
# read data
nz <- readr::read_csv2(url("https://raw.githubusercontent.com/go-bayes/psych-447/main/data/nz/nz.csv"))

# make characters into factors, note the "mutate_if" command
nz <-nz %>%
  dplyr::mutate_if(is.character, factor)

# make a date variable
nz <- nz %>%
  dplyr::mutate(date = make_date(year = 2009, month = 6, day = 30) + TSCORE) %>%
   dplyr::mutate_if(is.character, factor) %>%
  select(-c(SWB.Kessler01,SWB.Kessler02,SWB.Kessler03,SWB.Kessler04,SWB.Kessler05,
            SWB.Kessler06))%>%
  dplyr::mutate(Wave = as.factor(Wave))
```



### Q1a. Warmup

Make a new dataframe from the `nz` dataframe. Call it `df`

### Solution

```{r}
df <- nz
```


### Q1b. Change class in all instances of a class 

Make all the hours variables into integers 


### Solution

This would be the quick way:

```{r}
head(df)
df <-df %>%
  dplyr::mutate(across(starts_with("Hours"), as.integer))
```

But more likely they'll do it by hand, which is fine:

```{r eval=FALSE}
df<- df%>%
  mutate(Hours.Exercise = as.integer(Hours.Exercise),
         Hours.Internet = as.integer(Hours.Internet))
```

And so on... 

### Q2a.  Scale, center, transform

Create a new indicator that standardises the `Pol.Orient` variable, create a new indicator that centers the `Pol.Orient` variable, create a new indicator that centres the `Age` variable in decade-long units. Do this in a single piped workflow. 

**Print the head of the data frame so that we can see your work **

### Solution
```{r eval=FALSE}
df1<-df %>%
  dplyr::select(Pol.Orient,
         Age) %>%
  dplyr::mutate(
    conservative_s = scale(Pol.Orient),
    conservative_c = scale(Pol.Orient, center = TRUE, scale = FALSE),
    age_decade_c = scale(Age, center = TRUE, scale = FALSE) / 10
  )
head(df1)
```


### Q2b Data wrangle

Select `Hour.Exercise` and filter Wave 2019.

### Solution 

Note that this won't.  We need to filter first because otherwise there would be no Wave column to select!

```{r eval = FALSE}
df$Wave
df%>%
  dplyr::select("Hours.Exercise")%>%
  dplyr::filter(Wave == 2019)
```


Like so:

```{r}
df%>%
  dplyr::filter(Wave == 2019) %>%
  dplyr::select("Hours.Exercise")%>%
  head()
```


### Q3. Working with dates

What are the maximum number of responses for a single day in 2018 and the maximum number of responses for a single day in 2019? 


#### Solution

This is how I'd do it:

```{r}
nz %>%
  group_by(date, Wave)%>%
  count() %>%
  group_by(Wave)%>%
  dplyr::filter(n == max(n))
```


Here's a longer approach:

```{r}
# 2018 take the first
nz %>%
  group_by(date, Wave) %>%
  count() %>%
  filter(Wave == 2018) %>%
  arrange(desc(n)) %>%
  head()

# 2019  take the first
nz %>%
  group_by(date, Wave) %>%
  count() %>%
  filter(Wave == 2019) %>%
  arrange(desc(n))%>%
  head()
```


### Q4 Caculating dates and creating summaries

How many days are there between the date with the highest number of responses and the date with the second highest number of responses? 

Bonus: Calculate difference between the number of responses on the highest response date and second highest response date.


### Solution

Here is an approach:

```{r}  
ts<-nz %>%
  group_by(date)%>%
  count() %>%
  arrange(desc(n)) %>%
  as_tibble() %>%
  slice(c(1:2)) %>%
  mutate(diff_days = date - lag(date),
         diff_n = n - lag(n))
ts
```



### Q5. Working with date intervals

Suppose you were born on Dec 25, 1995 at 5.02:22 am Calculate your age in months on March 20,2021, at 1:22:04pm. (Hint use the `lubridate` package. Look up the `interval` function).


### Solution

```{r}
library(lubridate)
alive <-lubridate::interval(ymd_hms("1995-12-25 05:02:22"), ymd_hms("2021-03-21 13:22:04"))

time_length(alive, "months")
```


### Q6 Create an ordered factor from numeric data

The `Religion.Church` variable contains responses to the question: "How many times each month do you attend church or religious service?"

Create factor with the following three levels:

1. People who attend church 0 times per month, 
2. People who attend church 1-3 times per month, 
3. People who attend church 4 or more times per month. 

Make sure to re-level the factor so that the ordinal ranking moves from lowest to highest.


###  Solution

```{r}
nz %>%
  dplyr::mutate(church_attendance_cats = cut(
    Religion.Church,
    breaks = c(-Inf, 0, 3.99, Inf),
    labels = c("zero", "less_4", "gr_4"),
    right = TRUE,
  )) %>% 
  group_by(church_attendance_cats) %>%
  count()
```

### Another solution

```{r}
nz %>%
  dplyr::select(Religion.Church) %>%
  dplyr::mutate(church_attendance_cats = as.factor(ifelse(
    Religion.Church == 0,
    "zero",
    ifelse(Religion.Church < 4, "less_4", "gr_4")
  ))) %>%
  group_by(church_attendance_cats) %>%
  count()
```

### Q7 Make a summary table

Using methods described in lecture 4, create a table for average hours of sleep by month in the `nz` dataset



### Solution


```{r}
library(kableExtra)
tabnz<-nz %>%
  select(Id, date, HLTH.SleepHours) %>%
  mutate(month = month(date, label = TRUE))%>%
  group_by(month) %>%
  summarise(
    average_sleep =  mean(HLTH.SleepHours, na.rm = TRUE),
    sd_sleep  =  sd(HLTH.SleepHours, na.rm = TRUE),
    n  = n()  ) 
tabnz%>%
  kbl(caption = "Distress by month") %>%
   kable_classic_2(c("striped", "hover"), full_width = TRUE)
```


### Q7. Make a summary graph

Graph the average hours of sleep by month including 95% confidence intervals

Briefly explain why some intervals are wider than others. 

### Solution
```{r, eval = FALSE, echo = FALSE}
nz %>%
  select(Id, date, HLTH.SleepHours) %>%
  mutate(month = month(date, label = TRUE))%>%
  group_by(month) %>%
  summarise(
    mn_sh =  mean(HLTH.SleepHours, na.rm = TRUE),
    sd_sh  =  sd(HLTH.SleepHours, na.rm = TRUE),
    n_sh  = n()  ) %>%
  mutate(
    se_sh  = sd_sh  / sqrt(n_sh ),
    lw_ci = mn_sh  - qt(1 - (0.05 / 2), n_sh  - 1) * se_sh ,
    up_ci = mn_sh  + qt(1 - (0.05 / 2), n_sh  - 1) * se_sh 
  ) %>%
  ggplot(., aes(x = month, y = mn_sh , colour = mn_sh )) +
  geom_errorbar(aes(ymin = lw_ci, ymax = up_ci), width = .1) +
  geom_point(size = 3)  +
  scale_y_continuous(limits = c(0,8)) + 
  theme_classic() + scale_fill_viridis_d()
```


Some intervals are wider than others because the number of responses by month varies, and a confidence interval divides by the number of cases.

```{r, eval = FALSE, echo = FALSE}
gnz <-nz %>%
  select(Id, date, HLTH.SleepHours,Wave) %>%
  mutate(month = month(date, label = TRUE))%>%
  group_by(month, Wave) %>%
  summarise(
    mn_sh =  mean(HLTH.SleepHours, na.rm = TRUE),
    sd_sh  =  sd(HLTH.SleepHours, na.rm = TRUE),
    n_sh  = n()  ) %>%
  mutate(
    se_sh  = sd_sh  / sqrt(n_sh ),
    lw_ci = mn_sh  - qt(1 - (0.05 / 2), n_sh  - 1) * se_sh ,
    up_ci = mn_sh  + qt(1 - (0.05 / 2), n_sh  - 1) * se_sh 
  )
```

Here we see high standard errors of the mean with low n's

```{r, eval = FALSE, echo = FALSE}
gnz%>%
  select(month,Wave,mn_sh,se_sh,n_sh)
```

We can graph the year-wise differences, focusing on April:

```{r, eval = FALSE, echo = FALSE}
gnz%>%
  ggplot(., aes(x = month, y = mn_sh , colour = factor(Wave) )) +
  geom_errorbar(aes(ymin = lw_ci, ymax = up_ci), width = .1) +
  geom_point(size = 3)  +
 # scale_y_continuous(limits = c(0,8)) + 
  theme_classic() + scale_colour_viridis_d()
```


### Q8. Correlation graph

Created a correlation graph for the items in the Kessler 6 scale

These are `FeelHopeless,FeelDepressed,FeelRestless,EverythingIsEffort,FeelWorthless,FeelNervous`

Hint you must transform the factors into integers.

What do you find most interesting about this plot? Explain.


### Solution

```{r}
k6<-nz%>%
  select(FeelHopeless,FeelDepressed,FeelRestless,EverythingIsEffort,
         FeelWorthless,FeelNervous, Id)%>%
  mutate_all(.,as.numeric) %>%
  mutate(Id= as.factor(Id))# make numeric for correlation plot

k6 %>%
  correlation::correlation( partial = FALSE, multilevel = TRUE ) %>%
  plot() + theme_gray()

```

I find it interesting that `Nervousness` and `Restlessness` are not correlated, or inversely correlated with `Depression.`

Setting the correlations to `partial` does not destroy this effect.

```{r}
k6 %>%
  correlation::correlation( partial = TRUE, multilevel = TRUE ) %>%
  plot() + theme_gray()
```


### Q9 Create a blank `papaja` report 

Include your your name, affiliation, contributors and r packages used in your analysis


### Q10

```{r dualfig, fig.cap="\\label{fig:dualfig}"}
library(patchwork)
library(ggplot2)
p1 <- qplot(mtcars$cyl, geom = "histogram") + labs(title = "this plot") + xlab("mt cycle")
p2 <- qplot(mtcars$disp, geom = "histogram")+ labs(title = "that plot")

p1/p2 + plot_annotation(title = "my title", tag_levels = 'a') + xlab("mt cycle") +  plot_layout(guides = 'collect')

```


