---
title: "Week 11 workbook and solutions"
description: |
author:
  - name: Joseph Bulbulia
    url: https://josephbulbulia.netlify.app
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0002-5861-2056
output:
  distill::distill_article:
    self_contained: false
    toc: true
    highlight: kate
    code_folding: true
---

Import the jittered NZAVS dataset

```{r setup, include = FALSE}
# setup
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  layout = "l-body-outset",
  fig.width= 12,
  fig.height= 10,
  collapse =TRUE,
  R.options = list(width = 60)
)
```

```{r}
### Libraries
library("tidyverse")
library("patchwork")
library("lubridate")
library("kableExtra")
library("gtsummary")
library("lubridate")
library("equatiomatic")
library("ggdag")
library("brms")
library("rstan")
library("rstanarm")
library("bayesplot")
library("easystats")
library("kableExtra")
library("broom")
library("tidybayes")
library("bmlm")
# rstan options
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores ())
theme_set(theme_classic())
```

```{r  nzdata, include = TRUE}
# read data

nz_0 <- as.data.frame(readr::read_csv2(
  url(
    "https://raw.githubusercontent.com/go-bayes/psych-447/main/data/nzj.csv"
  )
))

# to re-level kessler 6 variables
f <-
  c(
    "None Of The Time",
    "A Little Of The Time",
    "Some Of The Time",
    "Most Of The Time",
    "All Of The Time"
  )

# get data into shape
nz_cr <- nz_0 %>%
  dplyr::mutate_if(is.character, factor) %>%
  dplyr::select(
    -c(
      SWB.Kessler01,
      SWB.Kessler02,
      SWB.Kessler03,
      SWB.Kessler04,
      SWB.Kessler05,
      SWB.Kessler06
    )
  ) %>% 
  dplyr::mutate(Wave = as.factor(Wave)) %>%
  dplyr::mutate(FeelHopeless = forcats::fct_relevel(FeelHopeless, f)) %>%
  dplyr::mutate(FeelDepressed = forcats::fct_relevel(FeelDepressed, f)) %>%
  dplyr::mutate(FeelRestless = forcats::fct_relevel(FeelRestless, f)) %>%
  dplyr::mutate(EverythingIsEffort = forcats::fct_relevel(EverythingIsEffort, f)) %>%
  dplyr::mutate(FeelWorthless = forcats::fct_relevel(FeelWorthless, f)) %>%
  dplyr::mutate(FeelNervous = forcats::fct_relevel(FeelNervous, f)) %>%
  dplyr::mutate(Wave = as.factor(Wave)) %>%
  dplyr::mutate(male_id = as.factor(Male)) %>%
  dplyr::mutate(date = make_date(year = 2009, month = 6, day = 30) + TSCORE) %>%
  dplyr::mutate(
    FeelWorthless_int = as.integer(FeelWorthless),
    FeelNervous_int =  as.integer(FeelNervous),
    FeelHopeless_int =  as.integer(FeelHopeless),
    EverythingIsEffort_int =  as.integer(EverythingIsEffort),
    FeelRestless_int =  as.integer(FeelRestless),
    FeelDepressed_int =  as.integer(FeelDepressed),
    HLTH.Fatigue_int = as.integer(HLTH.Fatigue + 1)
  ) %>%
  dplyr::mutate( yearS = (TSCORE - min(TSCORE)/365) ) %>%
  dplyr::mutate(KESSLER6sum = as.integer(KESSLER6sum))

## if you do anything with covid (warning, such a model would be tricky)

ord_dates_class <- c("Baseline",
                     "PreCOVID",
                     "JanFeb",
                     "EarlyMarch",
                     "Lockdown",
                     "PostLockdown")

nzl <- nz_cr %>%
  dplyr::filter(YearMeasured == 1) %>%
  dplyr::mutate(yearS = (TSCORE - min(TSCORE)/365)) %>%
  dplyr::mutate(WSCORE = as.factor(WSCORE)) %>%
  dplyr::mutate(Covid_Timeline =
                  as.factor(ifelse(
                    TSCORE %in% 3896:3921,
                    # feb 29 - march 25th
                    "EarlyMarch",
                    ifelse(
                      TSCORE %in% 3922:3954,
                      "Lockdown",
                      #march 26- Mon 27 April 2020
                      ifelse(
                        TSCORE > 3954,
                        # after april 27th 20202
                        "PostLockdown",
                        ifelse(
                          TSCORE %in% 3842:3895,
                          # jan 6 to feb 28
                          "JanFeb",
                          ifelse(TSCORE %in% 3665:3841 &
                                   Wave == 2019,
                                 "PreCOVID",
                                 "Baseline"  # 3672 TSCORE or  20 July 2019))))))))
                          )
                        )
                      )
                    ))))%>%
  dplyr::mutate(Covid_Timeline = forcats::fct_relevel(Covid_Timeline, ord_dates_class))%>%
  dplyr::mutate(Id = factor(Id))

```

```{r include = FALSE}
# inspect data
dplyr::glimpse(nzl)
```

## Assessment: briefly answer three of the following five questions. 

1. What is the difference between prediction and causal inference: give an example of how regression may be use useful for each task. Make sure your explanation includes a DAG.

2. Explain "collider confounding" (or "collider bias"), and explain how collider bias can spoil inference. Make sure that your explanation includes a DAG.

3. Using the `nzl` dataset, select at least five demongraphic/ideological variables that might be related to an exposure variable and and outcome variable of your choice. Create a DAG and identify which variables you should include to obtain an unbiased estimate of the causal effect of your exposure variable on your outcome variable. Test your model and interpret the results.

4. Write a DAG for one of your previous workbook regression analyses; Explain your DAG.  Redo your analysis to obtain an unbiased causal inference (conditional on your DAG). Compare the results of your current model with the results of your previous model and note any differences in interpretation. 

5. Examine the causal assumptions of a psychological theory for which you have data. Test the causal assumptions of this model. Make sure to draw a DAG. Consider simulating data if/where relevant. 


<!-- 6.  Import this dataset:  -->

<!-- According to a psychological theory that one of you is working on, "prototypicality of M훮ori group membership moderates the effect of perceived discrimination on ethnic identity in M훮ori." -->

<!-- Using the following three variables are from *Wave 4* of the NZAVS-jitter dataset, assess the causal claims of this theory, noting ways in which the theory might be improved.  Make sure to include a your DAG and make sure to clarify your DAG's assumptions. [Here, there is no need to draw on previous psychological theories.] -->

<!-- -   T4.MAORI.GME (M훮ori Group membership and evaluation) -->

<!-- -   T4.MAORI.VEM (M훮ori Perceived Appearance) -->

<!-- -   Perc.Discrim.T04 (Perceived discrimination - ethnicity) -->

<!--     -   I Feel that I am often discriminated against because of my ethnicity. -->
<!--     -   I feel that I am often discriminated against because of my ethnicity. -->



## Marking criteria

1.  Clarity, accuracy, initiative, and organisation in the descriptive component of your work.

2.  Clarity, accuracy, initiative, and organisation in your DAG and, if relevant, your regression model(s).

3.  Accuracy and insight in the interpretation of DAG, and if relevant, your regression model(s).




## Supplement: NZ-jitter information: 

Link to the NZAVS data dictionary is [here](https://dl.dropboxusercontent.com/s/but0pz9h4jdpy6f/NZAVS-Technical-Documents-e02-Data-Dictionary.xlsx?dl=0)

Link to questions only [here](https://dl.dropboxusercontent.com/s/cv256sbjal4ucvo/nzavs-data-dictionary.htm?dl=0)
