---
title: "Week 2 Lab"
description: |
author:
  - name: Joseph Bulbulia
    url: https://josephbulbulia.netlify.app
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0002-5861-2056
  - name: Johannes Karl 
    url: https://johannes-karl.com
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0001-5166-0728
output:
  distill::distill_article:
    self_contained: false
    toc: true
---


## Workflow

If you haven't already, set a folder for your R projects. Choose a location on our Hard drive. It is best *not* to use a shared folder (iCloud, Dropbox, Google Drive) because this can sometime cause problems. GitHub will be your place to store your code. 

Mine is called "GIT".


Using the methods you learned last week:

### Create Repos

  1. Create a repository for 447-journals (if you don't have one already)
  2. Create a repository for your 447-workbooks. 
  3. You should initialise a Readme statement and a .gitignore document (make sure to click the R .gitignore) 


### Create R projects

  1. Create R-projects in each of these repositories.
  2. In your workbooks folder, using the `+ New Folder` command, create a folder called `data` (or similar) This is where you'll store your data.
  4. Create a file called `figs` (where you'll stor your figures)
  4. Create a folder called `workbooks`, or similar this is where you'll store your weekly workbooks. 
  
### Workbook 2

  1. Open an R markdown document, and filling in your name
  2. The first code chunk is for your preferences. 


```{r setup, include=TRUE, eval = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



  3. Find the keyboard shortcuts menu. 
  4. Memorise the keyboard shortcut for insert code chunk which is: ⌥⌘I
  5. Use the mac to insert a code chunk. 
  6. Memorise the shortcut for evaluate which is ⌘-return^[there are several flavours of evaluation, COMMAND-RETURN is generically useful.
  

```{r, echo = TRUE}
# Use r to calculate 10 calculations
3*3
```
  
```{r}
4+5
```

```{r}
3^3
```

```{r}
factorial(10)/2
```

Use assignment for a calculation

```{r}
a <- 9
b <-2
a + b
(a * 2)/b
sqrt(a) + b
```


So you can see how to use R as a calculator

## Data types

What is a factor? 

What is an number?

What is an integer?

What is a character? 


## Inspecting data

1. Use the `str` command to describe the `iris dataset
2. Create a histogram for `Petal.Length`
3. Create a histogram for Petal.Length

4. What are the advantages and disadvantages of creating more breaks in the Petal.Length indicator? 

```{r}
hist(iris$Petal.Length, breaks = 100)
```

## Extra

In class we wrote the function

```{r}
go_bayes_factorial <- function (x) {
  # check is the number is negative, positive or zero
  if (x  < 0) {
    print("not even Ashley Bloomfield could make a factorial for a negative number")
  } else if (x == 0) {
    print("the factorial of zero is defined as 1")
  } else {
    for (i in 1:x)
      y <- 1
    for (i in 1:x) {
      y <- y * ((1:x)[i])
    }
    print(y)
  }
}
```


Note that this gets the r factorial function correct for whole integers

```{r}
go_bayes_factorial(7) == factorial(7)
```

However, what is happening here? 

```{r}
go_bayes_factorial(7.1) == factorial(7.1)
```

Can you rewrite the go_bayes_factorial function so that it yeilds the R function? 

```{r}
go_bayes_factorial2 <- function (x) {
  # check is the number is negative, positive or zero
  if (x  < 0) {
    print("not even Ashley Bloomfield could make a factorial for a negative number")
  } else if (x == 0) {
    print("the factorial of zero is defined as 1")
  } else {
    for (i in 1:x)
      y <- 1
    for (i in 1:x) {
      y <- y * ((1:x)[i])
    }
    print(y)
  }
}
```



