---
title: "Week 2 Lab"
description: |
author:
  - name: Joseph Bulbulia
    url: https://josephbulbulia.netlify.app
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0002-5861-2056
  - name: Johannes Karl 
    url: https://johannes-karl.com
    affiliation: Victoria University of Wellington
    affiliation_url: https://www.wgtn.ac.nz
    orcid_id: 0000-0001-5166-0728
output:
  distill::distill_article:
    self_contained: false
    toc: true
---


## Workflow

If you haven't already, set a folder for your R projects. Choose a location on our Hard drive. It is best *not* to use a shared folder (iCloud, Dropbox, Google Drive) because this can sometime cause problems. GitHub will be your place to store your code. 

Mine is called "GIT".


Using the methods you learned last week:

### Create Repos (if you have not already)

  1. Create a repository for 447-journals (if you don't have one already)
  2. Create a repository for your 447-workbooks. 
  3. You should initialise a Readme statement and a .gitignore document (make sure to click the R .gitignore) 


### Create R projects

  1. Create R-projects in each of these repositories.
  2. In your workbooks folder, using the `+ New Folder` command, create a folder called `data` (or similar) This is where you'll store your data.
  4. Create a file called `figs` (where you'll stor your figures)
  5 Create a folder called `workbooks`, or similar this is where you'll store your weekly workbooks. 
  
### Workbook 2

  1. Open an R markdown document, and filling in your name
  2. The first code chunk is for your preferences. 


```{r setup, include=TRUE, eval = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  3. Find the keyboard shortcuts menu. 
  4. Memorise the keyboard shortcut for insert code chunk which is: ⌥⌘I
  5. Use the mac to insert a code chunk. 
  6. Memorise the shortcut for evaluate which is ⌘-return^  use for all tasks in this workbook. 
  7. Use R to do 10 mathematical calculations 
  
```{r}
4+5
```

```{r}
3^3
```

Show your work 



  8. Using R, find the square root of 324
  9. In the `iris` dataset, find the flower with the longest Sepal.Length. (hint, use the `max` function.)  Show the column details for this flower only. 
  10. In the `iris` dataset, find the flower with the shortest Sepal.Length. (hint, use the `min` function.)
  11. Calculate the difference bewteen the flower with the longest and shortest Sepal.Length
  12. Make a 7 column  x  100 row dataframe using `c`, `:`, `dataframe`
  13. Rename the columns of the dataframe using the names of snow white dwarfs, or any names.
  14. Using the dataset `woman`, write a linear model for `height ~ weight`
  15. Create a table and coefficient plot for this model
  16. Using `ggeffects`, create a prediction plot for this model.

<aside>
Next week we'll be creating plots like this
```{r}
library(ggplot2)
ggplot2::ggplot(data = women) + 
  geom_point(mapping = aes(x = height, y = weight)) + geom_smooth(method = lm, aes(x = height, y = weight)) + theme_classic() 
```
<\aside>


  17.  Explain what is being calculated here

```{r}
sum(women$weight > 140) / length(women$weight)
```
   
  18. Calculate the proportion of women who are over the mean weight of women

```{r eval = FALSE, echo = FALSE}
mw <- mean(women$weight)
sum(women$weight > mw)/ length(women$weight)
```
  
```{r eval = FALSE, echo = FALSE}
mh<-mean(women$height)
sum(women$height > mh)/ length(women$height)
```


  19. What are the advantages and disadvantages of creating more breaks in the Petal.Length indicator?  Clarify your answer in a paragraph. 

```{r}
hist(iris$Petal.Length)
```

```{r}
hist(iris$Petal.Length, breaks = 100)
```

  20. Spot the error in this code
  
```{r eval = FALSE, echo = FALSE}
mh<-mean(women$height)
sum(women$weight > mh)/ length(women$height)
```

  21. Reorder the columns of the `woman` dataset so that weight comes before height. Then rename the columns "w" and "h". 

```{r, eval = FALSE, echo = FALSE}
head(women)
dfa <-cbind.data.frame(women$weight,women$height)
names(dfa)[]<-c("w","h")

```

  22. 
  
  
  
```{r echo = FALSE, eval = FALSE}
library(tidyverse)
n <- 100
yourtibble <-tibble(
  id = 1:n, 
  weight = rnorm(n, 80,20), 
  height = rnorm(n, weight*2, 10)
)

testdata1<-data.frame(yourtibble)
write_csv(testdata1,here::here("data", "testdata1.csv"))


saveRDS(yourtibble, here::here("data", "yourtibble.RDS"))
```



```{r}
library (readr)
testdata<- url("https://github.com/go-bayes/psych-447/blob/main/data/testdata1.csv")
head(testdata)
```




